PM Mungbean meta-analysis

### 2020-06-22 01_Meta-analysis_experiments.Rmd

This meta-analysis uses data collected by a number of researchers from the Department of Agriculture and Primary Industries and the University of Southern
Queensland.
It is worth noting in particular Professor Malcolm Riley's and Dr Sue Thompson's effort in establishing and coordinating early field trials between 2010 and 2016.

@@

**2001 - 2004 trials**  
Trial data were obtained from a non-published manuscript.  
Ratings for powdery mildew were on a 1-5 rating scale where 1 = no sign of disease, 2 = 1-30% diseased leaf area, 3 = 31-65% diseased leaf area, 4 = 66-99% diseased leaf area and 5 = complete defoliation.  
Comparing these trials which have not used the 1-9 rating scale will be converted using a
percentage change.

**2013 Field trials**  
Trials from 2013 were all conducted by the [Northern Growers Alliance (NGA)](https://www.nga.org.au/).
The raw data were obtained from Lawrie Price.
NGA data were formatted to be compliant with the 1 - 9 rating scale used in all other experiments. Detailed code for the conversion can be found in the following vignette [191115_PMMB_DataWrangling_PM.Rmd](./vignette/191115_PMMB_DataWrangling_PM.html).

### 1 - 9 Powdery mildew rating scale

   1. No powdery mildew colonies observed on any plant

   2. Small colonies in lower 1/3 of canopy, up to 75% of plants affected

   3. Colonies in lower 1/2 of canopy, >75% of plants affected

   4. Colonies in lower 2/3 of canopy, up to 75% of plants affected

   5. Colonies in lower 2/3 of canopy, >75% of plants affected

   6. Colonies in lower 2/3 of canopy, 100% of plants affected

   7. Colonies in lower 2/3 of canopy of 100% of plants, some plants with colonies also in top 1/3 of canopy

   8. Colonies to top of plant with >75% of plants affected

   9. Colonies to top of plant with 100% of plants affected and heavy leaf drop

>>>>

This meta-analysis uses data collected by a number of researchers from the Department of Agriculture and Primary Industries and the University of Southern Queensland.
It is worth noting in particular Professor Malcolm Riley's and Dr. Sue Thompson's effort in establishing and coordinating early field trials between 2010 and 2016.

@@

**2001 - 2004 trials**

Trial data were obtained from a non-published manuscript.

### 1 - 5 powdery mildew rating scale

Ratings for powdery mildew were on a 1 - 5 rating scale where

   1 - No sign of disease

   2 - 1 - 30&nbsp;% diseased leaf area

   3 - 31 - 65&nbsp;% diseased leaf area

   4 - 66 - 99&nbsp;% diseased leaf area and

   5 - Complete defoliation

Comparing these trials which have not used the 1 - 9 rating scale will be converted using a percentage change.

**2013 Field trials**

Trials from 2013 were all conducted by the [Northern Growers Alliance (NGA)](https://www.nga.org.au/).
The raw data were obtained from Lawrie Price.
NGA data were formatted to be compliant with the 1 - 9 rating scale used in all other experiments.
Detailed code for the conversion can be found in the following vignette [191115_PMMB_DataWrangling_PM.Rmd](./vignette/191115_PMMB_DataWrangling_PM.html).

### 1 - 9 powdery mildew rating scale

   1 - No powdery mildew colonies observed on any plant

   2 - Small colonies in lower 1/3 of canopy, up to 75&nbsp;% of plants affected

   3 - Colonies in lower 1/2 of canopy, >75&nbsp;% of plants affected

   4 - Colonies in lower 2/3 of canopy, up to 7&nbsp;5% of plants affected

   5 - Colonies in lower 2/3 of canopy, >75&nbsp;% of plants affected

   6 - Colonies in lower 2/3 of canopy, 100&nbsp;% of plants affected

   7 - Colonies in lower 2/3 of canopy of 10&nbsp;0% of plants, some plants with colonies also in top 1/3 of canopy

   8 - Colonies to top of plant with >75&nbsp;% of plants affected

   9 - Colonies to top of plant with 100&nbsp;% of plants affected and heavy leaf drop

### 2020-06-22 02_Preliminary_analysis.Rmd

In most years, regardless of the age of the crop, powdery mildew establishes in March.

@@

let's tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship. 
Let's do the plot again for each trial where there are no missing values.  
Note: The plots are ordered by mean yield of the trial.

@@

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i,]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control", ],
          col = "red")

}

```
The plots from the AUPDS is similar to the AUDPC, however according to the linear model the AUDPC is a better, significant fit.

>>>>

In most years, regardless of the age of the crop, powdery mildew establishes in March, which is to be expected.
The disease is driven by weather conditions and the plant is vulnerable at any growth stage.

@@

Let's tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship.
Let's do the plot again for each trial where there are no missing values.
Note: The plots are ordered by mean yield of the trial.

@@

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_means[PM_MB_means$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_means$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_means$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

The plots from the AUPDS is similar to the AUDPC, however according to the linear model the AUDPC is a better, significant fit.
